### 2021-08-02 manuscript/manuscript.tex
version sent to Steve and Tony

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmc_template/bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage{graphicx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \def\includegraphic{}
% \def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\newcommand{\bm}[1]{{\textbf{#1}}}
\endlocaldefs

%\newcommand{\bm}[1]{{\textbf{#1}}}
%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Continuous estimates of heat emission at Mt. Ruapehu using
the Unscented Kalman Smoother}

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n1},                        % id's of article notes, if any
   email={y.behr@gns.cri.nz}   % email address
]{\inits{YB}\fnm{Yannik} \snm{Behr}}
\author[
   addressref={aff1},
   email={s.sherburn@gns.cri.nz}
]{\inits{SS}\fnm{Steve} \snm{Sherburn}}
\author[
   addressref={aff1},
   email={t.hurst@gns.cri.nz}
]{\inits{TH}\fnm{Tony} \snm{Hurst}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%
    \orgname{GNS Science},
  \street{114 Karetoto Rd},
  \postcode{3384}
  \city{Taupo},
  \cny{New Zealand}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{sample}
\kwd{article}
\kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section{Introduction}

The amount and rate of heat emitted by a volcano provides important clues about
volcanic processes at depth. For example, a sudden rise in the heat output rate
combined with increased microgravity signals may result from the emplacement of
fresh magma at shallow depth \cite{Brown1991}. A slow decrease in the heat
emission rate together with a high carbon-sulfur ratio of the emitted gas may
point to the formation of a hydrothermal seal.

Volcanic lakes help with estimating heat emissions. Often, the majority of a
volcano's heat and gas output passes through the crater lake. Crater lakes act thus
like a filter, integrating heat and gas input from the vents entering the lake.

Monitoring changes in temperature, water mass, and ion concentration in these
lakes can thus serve as proxies for changes in gas- and steam-release by the
hydrothermal system and, ultimately, the melt zone beneath the volcano.

Ruapehu Crater Lake (Te Wai A-Moe in Te Reo Maori and referred to as RCL from
hereon) on Mt. Ruapehu, an active andesitic stratovolcano in the center of New
Zealand's North Island (Figure \ref{overview}), is one such example with records
of observations dating back to the 19th century \cite{Friedlander1898}. Mt.
Ruapehu erupted at least 602 times since 1830 with many phreatic and
phreato-magmatic eruptions and two major magmatic episodes \citep{Scott2013}.
Although the latest magmatic event between June 1995 to November 1997 and a
subsequent dam collapse changed the shape and volume of the lake, it has been
largely unchanged since then and the lake's bathymetry is well known from times
when it was empty [Tony Hurst, personal communication].

First attempts to quantify the amount of heat entering the lake date back to
1966 \citep{Dibble1966}. Since 2016 water temperature and lake level have been
monitored continuously and water samples for ion concentration measurements have
been taken roughly every two to three months. This enables us to continuously
estimate heat input into the lake and, in principle, automatically detect
changes as soon as they occur.

Inference of the heat input is typically done through a mass and energy balance
calculation: the sum of mass and energy flow into and out of the lake has to
account for the observed changes in lake temperature, water level and ion
dilution \cite{Hurst1981, Hurst1991, Stevenson1992, Fournier2009, Scott1994}. At
RCL, and probably at most other volcanic lakes, this is an underdetermined
problem, that is the number of unknowns exceeds the number of constraints.
Hence, it is necessary to constrain some of the parameters in the mass and
energy balance model using prior knowledge. Further, some observations and input
parameters come with large uncertainties and are sampled irregularly.

In the following, we propose a probabilistic (Bayesian) inference approach which
naturally handles these difficulties and transparently propagates uncertainties of
observations and prior assumptions into posterior uncertainties of heat input.

\begin{figure}
\includegraphics[width=12 cm]{figures/overview.png}  
\caption{Image of Mt. Ruapehu's crater lake from 2008 (\textcopyright Karen
	Britten, GNS Science). The red O marks the location of the lake outlet. The
	location of Mt. Ruapehu is shown as a red triangle on the insert map.}
  \label{overview}
\end{figure}

\section{Probabilistic inference method}\label{Pim}

\subsection{Physical model}\label{phm}

Following \cite{Hurst1991} and \cite{Stevenson1992} the mass and energy balance
can be written in terms of three coupled ordinary differential equations. The
rate of lake water mass ($M$) change can be described as:
\begin{equation}\label{ode_M}
	\frac{dM}{dt} = \dot{M_s} - \dot{M_o} - \dot{M_e} + \dot{M_i}
\end{equation}

Where $\dot{M_s}$ is the rate of water inflow through melt and precipitation,
$\dot{M_o}$
is the outflow rate through seepage and overflow, $\dot{M_e}$ is the rate of mass
loss through evaporation and $\dot{M_i}$ is the rate of steam input through the vent
system beneath the lake.

The energy content of the lake can be described with the notion of
enthalpy, $H$. At constant pressure, enthalpy is defined as:
\begin{equation}
	H(T) = H(T_0) + \int_{T_0}^{T}c_pM\tau d\tau
\end{equation}

where $T$ is temperature and $c_p$ is the specific heat. For all following
equations we take $H(T_0=0~^{\circ}C)$ as our reference enthalpy and
$c_p=c_w=constant$, with $c_w$ the specific heat of water, such that the
enthalpy of lake water at temperature $T$ can be written as $H(T) = c_wMT$.
Because RCL is at an altitude of $\sim$~2500~m and big parts of its catchment
are covered by snow and ice all year round we assume the temperature of incoming
water to be at 0 $^{\circ}C$.

The rate of enthalpy change of the lake water can thus be written as:
\begin{equation}\label{ode_E}
    \frac{dH}{dt}=\frac{d(c_wTM)}{dt}=c_wT\dot{M} + c_wM\dot{T} =
	-\dot{Q_e} - c_wT\dot{M_o} + \dot{Q_{eff}} + \dot{Q_r}
\end{equation}

where $T$ is the lake water temperature, $\dot{Q_r}$ is the rate of energy gain
due to solar irradiation (short-wavelength radiation), and $c_wT\dot{M_o}$ is
the rate of energy loss through outflow/seepage. $\dot{Q_e}$ is the rate of
surface energy losses due to evaporation (forced and free convection), sensible
heat, and long-wavelength radiation (net loss). Several empirical and
theoretical equations exist to describe the surface losses and we refer the
reader to previous publications for more detailed discussions
\cite{Stevenson1992, hurstCraterLakeEnergy2015}. There is no consensus on the
best quantitative model of evaporation from warm lakes and, in fact, the most
suited equation may depend on the location of and available observations from
the lake. Appendix \ref{A} describes the equations used in this study in detail
which include the recent development on forced (i.e. wind-induced) convection
\cite{sartoriCriticalReviewEquations2000}.

$\dot{Q_{eff}}$ is the effective rate of energy gain from volcanic sources after
heating incoming water:
\begin{equation}
	\dot{Q_{eff}}=\dot{Q_i} - c_wT\dot{M_s}
\end{equation}
with $\dot{Q_i}$ the rate of energy input from volcanic sources.

Following from Equation \ref{ode_E} the temperature change $\frac{dT}{dt}$ can then be written as:
\begin{equation}\label{ode_T}
	\frac{dT}{dt}=\frac{1}{c_wM}\left(-\dot{Q_e} - c_wT\dot{M_o} + \dot{Q_{eff}} + \dot{Q_r}\right )
	-\frac{T}{M}\dot{M}
\end{equation}

Finally, the change in ion concentration of, for example, $Mg^{2+}$ can be expressed as:
\begin{equation}\label{ode_X}
	\frac{dX}{dt}=-\dot{M_o}\frac{X}{M}
\end{equation}

Where $X$ is the total ion amount in the lake. This equation is only true under
the assumption that $X$ only changes due to dilution and there is no re-supply
between eruptive episodes. Equation \ref{ode_X} gives an estimate of the average
outflow over periods on the order of months but does not capture shorter term
variations in the outflow.

In a sensitivity analysis for $\dot{T}$ (Figure \ref{sens}) we can see a linear
relationship with respect to $\dot{Q_i}$ and $\dot{M_o}$ and a mildly
non-linear relationship with respect to $\dot{Q_i}$ and current season
(note that this assumes southern hemisphere seasons). The relationship with respect
to $T$ and windspeed, however, is strongly non-linear. This is because $\dot{Q_e}$
and $\dot{M_e}$ are non-linear functions of windspeed and $\dot{Q_i}$. This further 
complicates the inference of $\dot{Q_i}$.

\begin{figure}
    \includegraphics[width=12 cm]{figures/sensitivity_analysis.png}  
    \caption{Sensitivity analysis of $\dot{T}$ with respect to $\dot{Q_i}$ 
             and $\dot{M_o}$ (top panel), $\dot{Q_i}$ and season 
             (center panel), and $T$ and windspeed (bottom panel). The model
              parameters that have been held fixed are described in each 
              panel's title. Warmer colors indicate a temperature increase
              ($\dot{T} > 0$) and cooler colors a temperature decrease
              ($\dot{T} < 0$)}
      \label{sens}
\end{figure}

\subsection{Inference}\label{inf}

The objective of the inference problem is to estimate the rate at which heat
is entering the lake from the hydrothermal system below ($\dot{Q_i}$).
To constrain this we use the currently available observations of $T$, $M$,
$X$, and $\dot{M_o}$. Putting it more concisely, we are looking for

\begin{equation}\label{bsm}
    p(\bm{x}_k|\bm{y}_{1:t})
\end{equation}

where $\bm{y}_{1:t}$ is the time series of $m$ observations and $x_k$ comprises
the $n$ model parameters described in Section \ref{phm} at time step $k$. Note
that $x_k$ also includes $T$, $M$, $\dot{M_o}$, and $X$. We distinguish here
between the state of the lake and observations of this state
($T$, $M$, $X$, $\dot{M_o}$) which are to some degree uncertain.

Solving Equation \ref{bsm} using standard strategies like Markov Chain Monte
Carlo is computationally very expensive as the number of variables grows
exponentially with the length of the timeseries. An alternative are Bayesian
Filtering and Smoothing algorithms of which the Kalman Filter and Smoother are
the closed form solution for linear models \cite{Kalman1960, Rauch1965,
sarkkaBayesianFilteringSmoothing2013}.

Equation \ref{bsm} is the conditional probability of $\bm{x}_k$ given all
observations until time $t$. We assume that the states of the lake form a
Markov chain, meaning that the state of the lake at time step $k$, $\bm{x}_k$,
only depends on the state at adjacent time steps:

\begin{equation}\label{markov1}
    p(\bm{x}_k|\bm{x}_{1:k-1}, \bm{y}_{1:k-1}) = p(\bm{x}_k|\bm{x}_{k-1})
\end{equation}

and

\begin{equation}\label{markov2}
    p(\bm{x}_{k-1}|\bm{x}_{k:t}, \bm{y}_{k:t}) = p(\bm{x}_{k-1}|\bm{x}_{k})
\end{equation}

With this assumption and using \textit{Bayes' rule}, Equation \ref{bsm} can be
rewritten \cite{sarkkaBayesianFilteringSmoothing2013} as:

\begin{equation}\label{bsm1}
p(\bm{x}_k|\bm{y}_{1:t}) = p(\bm{x}_k|\bm{y}_{1:k})
    \int\left[\frac{p(\bm{x}_{k+1}|\bm{x}_k)p(\bm{x}_{k+1}|\bm{y}_{1:t})}{p(\bm{x}_{k+1}|\bm{y}_{1:k})}\right]d\bm{x}_{k+1}
\end{equation}

with

\begin{eqnarray}\label{bfilt1}
p(\bm{x}_{k+1}|\bm{y}_{1:k}) & = & \int p(\bm{x}_{k+1}|\bm{x}_k)p(\bm{x}_k|\bm{y}_{1:k})d\bm{x}_k \\
        \label{bfilt}
    p(\bm{x}_k|\bm{y}_{1:k}) & = & \frac{1}{z_k} p(\bm{y}_k|\bm{x}_k)p(\bm{x}_k|\bm{y}_{1:k-1}) \\
        \label{zetk}
    z_k & = & \int p(\bm{y}_k|\bm{x}_k) p(\bm{x}_k| \bm{y}_{1:k-1}) d\bm{x}_k  
\end{eqnarray}

using the \textit{Chapman-Kolmogorov} equation
\cite{sarkkaBayesianFilteringSmoothing2013} we can write:

\begin{equation}
p(\bm{x}_k|\bm{y}_{1:k-1}) = \int p(\bm{x}_k|\bm{x}_{k-1})p(\bm{x}_{k-1}|\bm{y}_{1:k-1})d\bm{x}_{k-1} 
\end{equation}

Equations \ref{bsm1} to \ref{zetk} describe a recursive way to compute Equation
\ref{bsm} in which we only need to know $\bm{x}_1$,
$p(\bm{x}_k|\bm{x}_{k-1})$, and $p(\bm{y}_k|\bm{x}_k)$. Equation \ref{bsm1} and
\ref{bfilt} are also known as the Bayesian Smoothing and Bayesian Filtering
equation, respectively. If $p(\bm{x}_k|\bm{x}_{k-1})$ and
$p(\bm{y}_k|\bm{x}_k)$ are linear transformations with normally distributed
errors, the \textit{Kalman filter} \cite{Kalman1960} and Kalman smoother
\cite{Rauch1965} represent closed form solutions. In our case, these
transformations are:

\begin{eqnarray}
p(\bm{x}_k|\bm{x}_{k-1}) & = & f(\bm{x}_{k-1}) + \bm{q}_{k-1} \\
    p(\bm{y}_k|\bm{x}_k) & = & \bm{x}_k[1:m] + \bm{r}_k   
\end{eqnarray}

where $f(\bm{x}_{k-1})$ is the physical model described in Section \ref{phm}
propagated forward in time using a fourth-order Runge-Kutta scheme. If we 
assume the errors, $\bm{r}_k$ and $\bm{q}_{k}$, to be normally distributed
($\bm{q}_{k} \sim N(\bm{0},
\bm{Q}_{k})$, $\bm{r}_k \sim N(\bm{0}, \bm{R}_k)$), we can compute an
approximate solution to \ref{bsm1} using the \textit{Unscented Kalman Filter}
\citep{Merwe2004, sarkkaBayesianFilteringSmoothing2013}.

We estimate $\bm{R}_k$ from the daily variations in the observations but we do
not know $\bm{x}_1$ and $\bm{Q}_{k}$. Setting $\bm{x}_1$ to an arbitrary value
within physically reasonable bounds but large uncertainty works well in
practice. While $\bm{Q}_{k}$ can be formally optimised we treat it here as a
design parameter that we modify such that the observations are matched within
their error bounds and the inferred heat input rate varies smoothly. The latter
is motivated by the fact that the physical model described in \ref{phm} is not
able to model short-term transient changes in lake volume and temperature. These
can be caused by rain or increased meltwater influx and the underlying
assumption of a perfectly mixed lake typically does not apply in these
situations.

\section{Results}
\subsection{Synthetic test}\label{syn_test}

To evaluate the inference scheme we design a synthetic test, consisting of a
very large cylinder, the same volume as RCL. We keep the windspeed and surface
water inflow constant at 4.5 m/s and 10 kt/day, respectively. To model the
outflow we assume the outlet to be a pipe 0.2 $m^2$ in cross section and then
use Bernoulli's equation to model the water outflow based on the water level in
the cylinder above the pipe. We assume that the incoming heat follows a
combination of a third-order polynomial and a sinusoid function and that the
enthalpy of the incoming steam is 2.8 MJ/kg which corresponds to the enthalpy of
saturated steam at hydrostatic pressures similar to those at the bottom of the
lake \cite{Mayhew1978}. To the resulting synthetic observations we add
uncertainties similar to those estimated from real observations (see Section
\ref{realdata}).

Figure \ref{syn_example} shows the result of the probabilistic inference of
heat input compared to the true input of the synthetic model. The observations
are well fit within their error bounds but we see some descrepancies between
the original and the recovered heat input rates, especially for very steep
changes in the input rate. This is the result of preferring a smooth solution
as already mentioned in Section \ref{inf}.

It is noteworthy that the mass influx and outflux rates are recovered
quite accurately despite having no data in the former and very little data
in the latter case.

\begin{figure}
	\includegraphics[width=12 cm]{figures/synthetic_inversion_uks.png}  
    \caption{Inversion result for the synthetic test described in Section 
        \ref{syn_test}. Turquoise lines and markers represent hypothetical
        observations; red dashed lines are the inversion results and black
        dashed lines are the true input to the synthetic test.}
  \label{syn_example}
\end{figure}

\subsection{Real data}\label{realdata}
Temperature at RCL is measured by an integrated-circuit temperature device
(Texas Instruments LM35) located roughly 1.9 m below the lake's current
overflow level. The water level sensor is a hydrostatic pressure sensor (First
Sensor KTE/KTU/KTW6000\ldots CS Series) and it is co-located with the
temperature sensor. Both sensors have a sampling interval of 15 minutes and
transmit data several times a day via satellite to GNS Science, which monitors
New Zealand's geohazards through its GeoNet project (www.geonet.org.nz
 last accessed 16/07/2021).
Water samples are taken 8-10 times a year from which subsequently, among many
other analytes, concentrations of $Mg^{++}$ and $Cl^-$ are determined at GNS
Science's Geothermal Analytical Laboratory. We will focus here on $Mg^{++}$
concentration as it shows less deviation from the assumption of constant
dilution than $Cl^-$.

To estimate uncertainties of temperature and water level observations, we 
compute their daily averages and standard deviations. Uncertainties of $Mg^{++}$
concentrations are estimated directly when two samples were taken on the same day.
Otherwise, their uncertainty is taken to be the median of directly estimated
uncertainties.

$\dot{M_o}$ is measured once or twice per year and we further know from field
observations that the lake is not overflowing beneath a certain water level. We
fit a sigmoid function to these observations assuming a maximum outflow rate of
250 l/s. We assume the uncertainty in $\dot{M_o}$ as the 95 percent confidence
interval of the curve-fitting.

As there is no permanent weather station close to RCL we assume the wind speed
to be 5 m/s. This was the average windspeed derived from a temporary deployment
of a weather buoy on RCL and adjacent permanent weather stations
\cite{hurstUseWeatherBuoy2012}.

Figure \ref{inference_result} shows $\dot{Q_i}$ for RCL for lake measurements
between 4 March 2016 and 7 July 2021. During this time the heat input rate
reached a maximum of 753 MW and was on average 165 MW. The average standard
deviation was 34 MW and generally increased for lower values of $\dot{Q_i}$. To
put this in context, currently installed geothermal power plants in New Zealand
have a combined output of $\sim$~1000MW.

\begin{figure}
 	\includegraphics[width=13 cm]{figures/uks_real_data.png}  
 	\caption{Inference results for the heat input rate $\dot{Q_i}$ at RCL
             between 4 March 2016 and 7 July 2021. The solid line shows the
             marginal probability and the uncertainty is displayed as shaded
             region.}
\label{inference_result}
\end{figure}

\section{Discussion and conclusion}
Inferring the heat input rate into a crater lake is an important part in
estimating the total energy budget of a volcanic system \cite{Brown1989a}.
Similar to a volcano's gas budget, it can either indicate migration of magma or
changes in permeability within the vent system.

Figure \ref{psd} shows the power spectral density (PSD) for the marginal heat
input rate shown in Figure \ref{inference_result}. It was computed using the
multi taper method \cite{prietoComputersGeosciencesFortran2009} which provides
not only an estimate of the PSD but also its uncertainty. The PSD shows that the
dominant periods in the heat input rate lie between $\sim$100 and $\sim$250
days. Whether these periods are due to cyclic processes in the magmatic system,
the hydrothermal system, or a combination of both requires further modelling of
these processes. Many studies have demonstrated the value of non-linear Kalman
Filters for inference and data-assimilation with non-linear, high-dimension
numerical models \cite{White2018a}. While beyond the scope of this study, the
next logical step, therefore, is to include numerical models of magma degassing
and hydrothermal fluid flow to model the heat input rate, volume, and enthalpy
of the steam entering the lake.

\begin{figure}
    \includegraphics[width=12 cm]{figures/psd_analysis.png}  
 	\caption{Power spectral density of the heat input rate 
             (Figure \ref{inference_result}) shown as a solid line
             with uncertainty displayed as shaded region.}
    \label{psd}
\end{figure}

Irrespective of the process that is responsible for the periodicity, it appears
that it has not changed significantly over the past five years. This may
suggest that the volcano has been in a dynamic equilibrium during this time.
This is further supported by the cumulative heat input (Figure \ref{cumheat}).
It shows a slight increase in heat input in mid-2017 but a fairly constant rate
since then. Of course this does not in any way forecast the volcano's future
behaviour but it may serve as a useful baseline to detect changes in the
volcano's state more easily.

\begin{figure}
    \includegraphics[width=12 cm]{figures/cumulative_heat.png}
    \caption{Cumulative heat input into RCL (solid line) and its 
    uncertainty (shaded region).}
    \label{cumheat}
\end{figure}

Previous studies of RCL's mass and energy balance found that they required very
high values of enthalpy for the incoming steam to avoid violating physical
boundary conditions such as negative values for the inflow rate of meteoric
water \cite{hurstCraterLakeEnergy2015, Stevenson1992}. Their explanation was the
existence of a heat pipe beneath the lake where gas and fluid phases coexist. As
the rising gas enters the lake relatively cool lake water enters the vent
leading to an overall lower addition of mass to the lake and resulting in an
enthalpy value that is double than expected for saturated steam at hydrostatic
pressures likely to be found at the bottom of the lake. By including
uncertainties in our observations we no longer require these high enthalpy
values to satisfy the physical constraints. Numerical modelling of hydrothermal
processes may also shed further light on the heating processes and whether or
not a heat pipe is a plausible mechanism.

In conclusion, the combination of the mass and energy model, developed in
previous studies \cite{hurstCraterLakeEnergy2015, Stevenson1992}, with the
probabilistic inference method we developed here provides a powerful method to
continuously infer the heat input rate into RCL. It allows us to combine
disparate data streams, quantify our beliefs in the results and thereby augment
our monitoring capabilities by keeping a finger on Mt. Ruapehu's energy budget.

\appendix
\section{Surface losses}\label{A}
\subsection{long-wavelength radiation}
The net loss through long-wavelength radiation can be written as:
\begin{equation}
    Q_{rad} = A\sigma(\epsilon_w T_s^4 - \epsilon_a T_a^4)
\end{equation}

where
\begin{table}[h!]
\begin{tabular}{cp{8cm}}
    A & is the surface of the lake (in $m^2$) \\
    $\sigma$ & is the Stefan-Boltzmann coefficient ($5.67e^{-8} \frac{W}{m^2 K^4}$) \\
    $\epsilon_w$ & is the lake emissivity \\
    $T_s$ & is the water surface temperature (here assumed to be 1 $^{\circ}C$ less than the
    measured temperature) \\
    $\epsilon_a$ & is the effective atmospheric emissivity \\
    $T_a$ & is the effective air temperature (here assumed to be 0.9 $^{\circ}C$) \\ 
\end{tabular}
\end{table}

\subsection{Evaporation}
As proposed by \cite{hurstCraterLakeEnergy2015} we use the equation of
\cite{adamsEvaporationHeatedWater1990} but replace the term for forced
convection by that of \cite{sartoriCriticalReviewEquations2000}:

\begin{equation}
    Q_{evap} = [(2.2(T_{sv} - T_{av})^{1/3}(e_s - e_a))^2 + 
    (L(0.00407u^{0.8}X_0^{-0.2})(e_s - e_a)/P_a)^2]^{1/2}
\end{equation}

where
\begin{table}[h!]
\begin{tabular}{cp{8cm}}
    $T_{sv}$, $T_{av}$ & are the virtual surface and air temperatures respectively.
    This correction accounts for the extra buoyancy of water vapour compared to air
    (see Equation \ref{vtemp} for details) \\
    $e_s$, $e_a$ & are the saturation vapour pressure and the ambient air vapour
    pressure respectively \\
    $L$ & is the latent heat of evaporation \\
    $u$ & is the wind speed \\
    $X_0$ & is the fetch, or characteristic length, of the lake\\
    $P_a$ & is the atmospheric pressure \\
\end{tabular}
\end{table}

\begin{equation}\label{vtemp}
T_{xv} = \frac{T_x}{(1-0.378e_x/P_a)} \qquad \mathrm{for}\quad x = s\ \mathrm{or}\ a
\end{equation}

\subsection{Sensible heat}
This describes the heat loss due to conduction and convection above a hot lake
and according to \cite{Stevenson1992} the ratio between sensible heat,
$Q_{sh}$, and heat loss due to evaporation, $Q_{evap}$ can be written as:

\begin{equation}
    \frac{Q_{sh}}{Q_{evap}}=R_{sh}=\frac{\rho c_a}{L} \frac{(T_s - T_a)}{(e_s - e_a)}
    \underbrace{\left [ \frac{(T_s-T_a)}{(T_{sv} - T_{av})} \right ]^{1/3}}_A 
\end{equation}

where

\begin{table}[h!]
\begin{tabular}{cp{8cm}}
    $\rho$ & is the air density \\
    $c_a$ & is the specific heat of air\\
\end{tabular}
\end{table}

We decided to ignore term A as it tends to be very close to 1.

\subsection{Putting it all together}
The final equation to compute surface heat losses is then:
\begin{equation}
    Q_e = Q_{rad} + Q_{evap}(1+R_{sh})
\end{equation}

and the mass-loss due to evaporation is:
\begin{equation}
    M_e = Q_{evap}/L
\end{equation}

%\nocite{oreg,schn,pond,smith,marg,hunn,advi,koha,mouse}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
    Text for this section \ldots

\section*{Acknowledgements}
  Text for this section \ldots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{zotero}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

%% \section*{Figures}
%%   \begin{figure}[h!]
%%   \caption{\csentence{Sample figure title.}
%%       A short description of the figure content
%%       should go here.}
%%       \end{figure}
%% 
%% \begin{figure}[h!]
%%   \caption{\csentence{Sample figure title.}
%%       Figure legend text.}
%%       \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
%% \section*{Tables}
%% \begin{table}[h!]
%% \caption{Sample table title. This is where the description of the table should go.}
%%       \begin{tabular}{cccc}
%%         \hline
%%            & B1  &B2   & B3\\ \hline
%%         A1 & 0.1 & 0.2 & 0.3\\
%%         A2 & ... & ..  & .\\
%%         A3 & ..  & .   & .\\ \hline
%%       \end{tabular}
%% \end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% \section*{Additional Files}
%%   \subsection*{Additional file 1 --- Sample additional file title}
%%     Additional file descriptions text (including details of how to
%%     view the file, if it is in a non-standard format or the file extension).  This might
%%     refer to a multi-page table or a figure.
%% 
%%   \subsection*{Additional file 2 --- Sample additional file title}
%%     Additional file descriptions text.

\end{backmatter}
\end{document}