### 2022-02-03 manuscript/manuscript.tex
Version 2 of the manuscript

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage{graphicx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \def\includegraphic{}
% \def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\newcommand{\bm}[1]{{\textbf{#1}}}
\endlocaldefs

%\newcommand{\bm}[1]{{\textbf{#1}}}
%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Continuous estimates of heat emission at Mt. Ruapehu using
the Unscented Kalman Smoother}

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   email={y.behr@gns.cri.nz}   % email address
]{\inits{YB}\fnm{Yannik} \snm{Behr}}
\author[
   addressref={aff1},
   email={s.sherburn@gns.cri.nz}
]{\inits{SS}\fnm{Steve} \snm{Sherburn}}
\author[
   addressref={aff1},
   email={t.hurst@gns.cri.nz}
]{\inits{TH}\fnm{Tony} \snm{Hurst}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%
    \orgname{GNS Science},
  \street{114 Karetoto Rd},
  \postcode{3384}
  \city{Taupo},
  \cny{New Zealand}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
    Volcanic lakes often capture a significant amount of volcanic heat emission
    and thus provide a unique opportunity to monitor changes inside the volcano.
    We present a Bayesian inversion method to automatically infer volcanic heat
    emission at the base of a volcanic lake from lake monitoring data using a
    non-linear Kalman Smoother. Our method accounts for the, in parts large,
    uncertainties in observations and the underlying physics-based model to
    generate probabilistic estimates of heat emission. We verify our results
    using a synthetic test case and then estimate the daily heat input rate into
    Mt. Ruapehu's Crater Lake, New Zealand, between 2016 and 2022. Time-frequency
    analysis of the heat input rate shows dominant periods of heating cycles
    ranged between 100 - 250 days. Some years are dominated by shorter cycles
    and others by longer cycles which potentially points to changes in the
    magmatic heat supply and the hydrothermal system.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{Crater lakes}
\kwd{Non-linear Kalman Smoother}
\kwd{Volcanic heat emission}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section{Introduction}

The amount and rate of heat emitted by a volcano provides important clues about
volcanic processes at depth. For example, a sudden rise in the heat output rate
combined with increased microgravity signals may result from the emplacement of
fresh magma at shallow depth \cite{Brown1991}. A slow decrease in the heat
emission rate together with a high carbon-sulfur ratio of the emitted gas may
point to the formation of a hydrothermal seal.

Volcanic lakes help with estimating heat emissions. Often, the majority of a
volcano's heat and gas output passes through the crater lake. Crater lakes act thus
like a filter, integrating heat and gas input from the vents entering the lake.

Monitoring changes in temperature, water mass, and ion concentration in these
lakes can, therefore, serve as proxies for changes in gas- and steam-release by
the hydrothermal system and, ultimately, the melt zone beneath the volcano.

Ruapehu Crater Lake (Te Wai A-Moe in Te Reo Maori and referred to as RCL from
hereon) on Mt. Ruapehu, an active andesitic stratovolcano in the center of New
Zealand's North Island (Figure \ref{overview}), is one such example with records
of observations dating back to the 19th century \cite{Friedlander1898}. Mt.
Ruapehu erupted at least 602 times since 1830 with many phreatic and
phreato-magmatic eruptions and two major magmatic episodes \citep{Scott2013}.
Although the latest magmatic event between June 1995 to November 1997 and a
subsequent dam collapse changed the shape and volume of the lake, it has been
largely unchanged since then and the lake's bathymetry is well known from times
when it was empty [Tony Hurst, personal communication].

First attempts to quantify the amount of heat entering the lake date back to
1966 \citep{Dibble1966}. Since 2016 water temperature and lake level have been
monitored continuously and water samples for ion concentration measurements have
been taken roughly every two to three months. This enables us to continuously
estimate heat input into the lake and, in principle, automatically detect
changes as soon as they occur.

Inference of the heat input is typically done through a mass and energy balance
calculation: the sum of mass and energy flow into and out of the lake has to
account for the observed changes in lake temperature, water level and ion
dilution \cite{Hurst1981, Hurst1991, Stevenson1992, Fournier2009, Scott1994}. At
RCL, and probably at most other volcanic lakes, this is an underdetermined
problem, that is the number of unknowns exceeds the number of constraints.
Hence, it is necessary to constrain some of the parameters in the mass and
energy balance model using prior knowledge. Further, some observations and input
parameters come with large uncertainties and are sampled irregularly.

In the following, we propose a probabilistic (Bayesian) inference approach which
naturally handles these difficulties and transparently propagates uncertainties of
observations and prior assumptions into posterior uncertainties of heat input.

\begin{figure}
\includegraphics[width=12 cm]{figures/overview.png}  
\caption{Image of Mt. Ruapehu's crater lake from 2008 (\textcopyright Karen
	Britten, GNS Science). The red O marks the location of the lake outlet. The
	location of Mt. Ruapehu is shown as a red triangle on the insert map.}
  \label{overview}
\end{figure}

\section{Probabilistic inference method}\label{Pim}

\subsection{Physical model}\label{phm}

The mass and energy balance model can be written in terms of three coupled ordinary
differential equations \cite{Hurst1991, Stevenson1992}. The rate of lake water
mass ($M$) change is described as:
\begin{equation}\label{ode_M}
	\frac{dM}{dt} = \dot{M_s} - \dot{M_o} - \dot{M_e} + \dot{M_i}
\end{equation}

Where $\dot{M_s}$ is the rate of water inflow through melt and precipitation,
$\dot{M_o}$
is the outflow rate through seepage and overflow, $\dot{M_e}$ is the rate of mass
loss through evaporation and $\dot{M_i}$ is the rate of steam input through the vent
system beneath the lake.

The energy content of the lake can be described with the notion of
enthalpy, $H$. At constant pressure, enthalpy is defined as:
\begin{equation}
	H(T) = H(T_0) + \int_{T_0}^{T}c_pM\tau d\tau
\end{equation}

where $T$ is temperature and $c_p$ is the specific heat. For all following
equations we take $H(T_0=0~^{\circ}C)$ as our reference enthalpy and
$c_p=c_w=constant$, with $c_w$ the specific heat of water, such that the
enthalpy of lake water at temperature $T$ can be written as $H(T) = c_wMT$.
Because RCL is at an altitude of $\sim$~2500~m and big parts of its catchment
are covered by snow and ice all year round we assume the temperature of incoming
water to be at 0 $^{\circ}C$.

The rate of enthalpy change of the lake water can thus be written as:
\begin{equation}\label{ode_E}
    \frac{dH}{dt}=\frac{d(c_wTM)}{dt}=c_wT\dot{M} + c_wM\dot{T} =
	-\dot{Q_e} - c_wT\dot{M_o} + \dot{Q_{eff}} + \dot{Q_r}
\end{equation}

where $T$ is the lake water temperature, $\dot{Q_r}$ is the rate of energy gain
due to solar irradiation (short-wavelength radiation), and $c_wT\dot{M_o}$ is
the rate of energy loss through outflow/seepage. $\dot{Q_e}$ is the rate of
surface energy losses due to evaporation (forced and free convection), sensible
heat, and long-wavelength radiation (net loss). Several empirical and
theoretical equations exist to describe the surface losses and we refer the
reader to previous publications for more detailed discussions
\cite{Stevenson1992, hurstCraterLakeEnergy2015}. There is no consensus on the
best quantitative model of evaporation from warm lakes and, in fact, the most
suited equation may depend on the location of and available observations from
the lake. Appendix \ref{A} describes the equations used in this study in detail
which include the recent development on forced (i.e. wind-induced) convection
\cite{sartoriCriticalReviewEquations2000}. It is important to note here that
$Q_e$ is a function of $T$ and the windspeed above the lake.

$\dot{Q_{eff}}$ is the effective rate of energy gain from volcanic sources after
heating incoming water:
\begin{equation}
	\dot{Q_{eff}}=\dot{Q_i} - c_wT\dot{M_s}
\end{equation}
with $\dot{Q_i}$ the rate of energy input from volcanic sources.

Following from Equation \ref{ode_E} the temperature change $\frac{dT}{dt}$ can then be written as:
\begin{equation}\label{ode_T}
	\frac{dT}{dt}=\frac{1}{c_wM}\left(-\dot{Q_e} - c_wT\dot{M_o} + \dot{Q_{eff}} + \dot{Q_r}\right )
	-\frac{T}{M}\dot{M}
\end{equation}

Finally, the change in ion concentration of, for example, $Mg^{2+}$ can be expressed as:
\begin{equation}\label{ode_X}
	\frac{dX}{dt}=-\dot{M_o}\frac{X}{M}
\end{equation}

Where $X$ is the total ion amount in the lake. This equation is only true under
the assumption that $X$ only changes due to dilution and there is no re-supply
between eruptive episodes. Equation \ref{ode_X} gives an estimate of the average
outflow over periods on the order of months but does not capture shorter term
variations in the outflow.

The main challenge in solving the mass and energy balance equations for
$\dot{Q_i}$ arises from the non-linear relationship between $Q_e$ and ${T}$
as shown in Figure \ref{sens}.

\begin{figure}
    \includegraphics[width=12 cm]{figures/sensitivity_analysis.png}  
    \caption{Surface energy loss ($Q_e$) with respect to lake water temperature
             ($T$). The model parameters that have been held fixed are shown in  
             the title.}
      \label{sens}
\end{figure}

\subsection{Inference}\label{inf}

The objective of the inference problem is to estimate the rate at which heat
is entering the lake from the hydrothermal system below ($\dot{Q_i}$).
To constrain this we use the currently available observations of $T$, $M$,
$X$, and $\dot{M_o}$. Putting it more concisely, we are looking for

\begin{equation}\label{bsm}
    p(\bm{x}_k|\bm{y}_{1:t})
\end{equation}

where $\bm{y}_{1:t}$ is the time series of $m$ observations and $x_k$ comprises
the $n$ model parameters described in Section \ref{phm} at time step $k$. Note
that $x_k$ also includes $T$, $M$, and $X$. We distinguish here between the
state of the lake and observations of this state ($T$, $M$, $X$) which are to
some degree uncertain. Equation \ref{bsm} can be seen as the probabilistic
formulation of the inverse problem inferring $\dot{Q_i}$ from available
observations.

Solving Equation \ref{bsm} using standard strategies like Markov Chain Monte
Carlo is computationally very expensive as the number of variables grows
exponentially with the length of the timeseries and the number of model
parameters. An alternative are Bayesian Filtering and Smoothing algorithms of
which the Kalman Filter and Smoother are the closed form solution for linear
models \cite{Kalman1960, Rauch1965, sarkkaBayesianFilteringSmoothing2013}.

Equation \ref{bsm} is the conditional probability of $\bm{x}_k$ given all
observations until time $t$. We assume that the states of the lake form a
Markov chain, meaning that the state of the lake at time step $k$, $\bm{x}_k$,
only depends on the state at adjacent time steps:

\begin{equation}\label{markov1}
    p(\bm{x}_k|\bm{x}_{1:k-1}, \bm{y}_{1:k-1}) = p(\bm{x}_k|\bm{x}_{k-1})
\end{equation}

and

\begin{equation}\label{markov2}
    p(\bm{x}_{k-1}|\bm{x}_{k:t}, \bm{y}_{k:t}) = p(\bm{x}_{k-1}|\bm{x}_{k})
\end{equation}

With this assumption and using \textit{Bayes' rule}, Equation \ref{bsm} can be
rewritten \cite{sarkkaBayesianFilteringSmoothing2013} as:

\begin{equation}\label{bsm1}
p(\bm{x}_k|\bm{y}_{1:t}) = p(\bm{x}_k|\bm{y}_{1:k})
    \int\left[\frac{p(\bm{x}_{k+1}|\bm{x}_k)p(\bm{x}_{k+1}|\bm{y}_{1:t})}{p(\bm{x}_{k+1}|\bm{y}_{1:k})}\right]d\bm{x}_{k+1}
\end{equation}

with

\begin{eqnarray}\label{bfilt1}
p(\bm{x}_{k+1}|\bm{y}_{1:k}) & = & \int p(\bm{x}_{k+1}|\bm{x}_k)p(\bm{x}_k|\bm{y}_{1:k})d\bm{x}_k \\
        \label{bfilt}
    p(\bm{x}_k|\bm{y}_{1:k}) & = & \frac{1}{z_k} p(\bm{y}_k|\bm{x}_k)p(\bm{x}_k|\bm{y}_{1:k-1}) \\
        \label{zetk}
    z_k & = & \int p(\bm{y}_k|\bm{x}_k) p(\bm{x}_k| \bm{y}_{1:k-1}) d\bm{x}_k  
\end{eqnarray}

using the \textit{Chapman-Kolmogorov} equation
\cite{sarkkaBayesianFilteringSmoothing2013} we can write:

\begin{equation}
p(\bm{x}_k|\bm{y}_{1:k-1}) = \int p(\bm{x}_k|\bm{x}_{k-1})p(\bm{x}_{k-1}|\bm{y}_{1:k-1})d\bm{x}_{k-1} 
\end{equation}

Equations \ref{bsm1} to \ref{zetk} describe a recursive way to compute Equation
\ref{bsm} in which we only need to know $\bm{x}_1$,
$p(\bm{x}_k|\bm{x}_{k-1})$, and $p(\bm{y}_k|\bm{x}_k)$. Equation \ref{bsm1} and
\ref{bfilt} are also known as the Bayesian Smoothing and Bayesian Filtering
equation, respectively. If $p(\bm{x}_k|\bm{x}_{k-1})$ and
$p(\bm{y}_k|\bm{x}_k)$ are linear transformations with normally distributed
errors, the \textit{Kalman filter} \cite{Kalman1960} and Kalman smoother
\cite{Rauch1965} represent closed form solutions. In our case, these
transformations are:

\begin{eqnarray}
p(\bm{x}_k|\bm{x}_{k-1}) & = & f(\bm{x}_{k-1}) + \bm{q}_{k-1} \\
    p(\bm{y}_k|\bm{x}_k) & = & \bm{x}_k[1:m] + \bm{r}_k   
\end{eqnarray}

where $f(\bm{x}_{k-1})$ is the non-linear physical model described in Section
\ref{phm} propagated forward in time using a fourth-order Runge-Kutta scheme. If
we assume the errors, $\bm{r}_k$ and $\bm{q}_{k}$, to be normally distributed
($\bm{q}_{k} \sim N(\bm{0}, \bm{Q}_{k})$, $\bm{r}_k \sim N(\bm{0}, \bm{R}_k)$),
we can compute an approximate solution to \ref{bsm1} using the \textit{Unscented
Kalman Smoother} \citep{Merwe2004, sarkkaBayesianFilteringSmoothing2013}, a
non-linear Kalman Smoother.

We estimate $\bm{R}_k$ from the daily variations in the observations but we do
not know $\bm{x}_1$ and $\bm{Q}_{k}$. Setting $\bm{x}_1$ to an arbitrary value
within physically reasonable bounds but large uncertainty works well in
practice. While $\bm{Q}_{k}$ can be formally optimised we treat it here as a
design parameter that we modify such that the observations are matched within
their error bounds and the inferred heat input rate varies smoothly. The latter
is motivated by the fact that the physical model described in \ref{phm} is not
able to model short-term transient changes in lake volume and temperature. These
can be caused by rain or increased meltwater influx and the underlying
assumption of a perfectly mixed lake typically does not apply in these
situations.

\section{Results}
\subsection{Synthetic test}\label{syn_test}

To evaluate the inference scheme we design a synthetic test, consisting of a
very large cylinder, the same volume as RCL. We keep the windspeed and surface
water inflow constant at 4.5 m/s and 10 kt/day, respectively. To model the
outflow we assume the outlet to be a pipe 0.2 $m^2$ in cross section and then
use Bernoulli's equation to model the water outflow based on the water level in
the cylinder above the pipe. We assume that the incoming heat follows a
combination of a third-order polynomial and a sinusoid function and that the
enthalpy of the incoming steam is 2.8 MJ/kg which corresponds to the enthalpy of
saturated steam at hydrostatic pressures similar to those at the bottom of the
lake \cite{Mayhew1978}. To the resulting synthetic observations we add
uncertainties similar to those estimated from real observations (see Section
\ref{realdata}). As with real data, average and standard deviation of all values
within a day are calculated. The standard deviation is taken as a proxy for the
uncertainty in observations ($\bm{R}_k$). To simulate the disparity between sampling
intervals of different data streams we randomly remove most of the observations
of $X$, $\dot{M_o}$, and $\dot{M_s}$. This dataset is then inverted using the
Unscented Kalman Smoother to recover the original heat input rate.

Figure \ref{syn_example} shows the result of the probabilistic inference of
heat input compared to the true input of the synthetic model. The observations
are well fit within their error bounds but we see some descrepancies between
the original and the recovered heat input rates, especially for very steep
changes in the input rate. This is the result of preferring a smooth solution
as already mentioned in Section \ref{inf}.

It is noteworthy that the mass outflux rates are recovered quite accurately
despite providing very little synthetic observations.

\begin{figure}
	\includegraphics[width=12 cm]{figures/synthetic_inversion_uks.png}  
    \caption{Inversion result for the synthetic test described in Section 
        \ref{syn_test}. Turquoise lines and markers represent synthetic 
        observations; red dashed lines are the inversion results and black
        dashed lines are the true input to the synthetic test. Error bars 
        represent $\pm$ 3 standard deviations.}
  \label{syn_example}
\end{figure}

\subsection{Real data}\label{realdata}
Temperature at RCL is measured by an integrated-circuit temperature device
(Texas Instruments LM35) located roughly 1.9 m below the lake's current
overflow level. The water level sensor is a hydrostatic pressure sensor (First
Sensor KTE/KTU/KTW6000\ldots CS Series) and it is co-located with the
temperature sensor. Both sensors have a sampling interval of 15 minutes and
transmit data several times a day via satellite to GNS Science, which monitors
New Zealand's geohazards through its GeoNet project \cite{GeoNetHome}.
Water samples are taken 8-10 times a year from which subsequently, among many
other analytes, concentrations of $Mg^{++}$ and $Cl^-$ are determined at GNS
Science's Geothermal Analytical Laboratory. We will focus here on $Mg^{++}$
concentration as it shows less deviation from the assumption of constant
dilution than $Cl^-$.

To estimate uncertainties of temperature and water level observations, we 
compute their daily averages and standard deviations. Uncertainties of $Mg^{++}$
concentrations are estimated directly when two samples were taken on the same day.
Otherwise, their uncertainty is taken to be the median of directly estimated
uncertainties.

$\dot{M_o}$ is measured once or twice per year and we further know from field
observations that the lake is not overflowing beneath a certain water level. We
fit a sigmoid function to these observations assuming a maximum outflow rate of
250 l/s. We assume the uncertainty in $\dot{M_o}$ as the 95 percent confidence
interval of the curve-fitting.

As there is no permanent weather station close to RCL we assume the wind speed
to be 5 m/s. This was the average windspeed derived from a temporary deployment
of a weather buoy on RCL and adjacent permanent weather stations
\cite{hurstUseWeatherBuoy2012}.

Figure \ref{inference_result}a shows $\dot{Q_i}$ for RCL for lake measurements
between 4 March 2016 and 1 February 2022. During this time the heat input rate
reached a maximum of 753 MW and was on average 165 MW. The average standard
deviation was 34 MW and generally increased for lower values of $\dot{Q_i}$. To
put this in context, currently installed geothermal power plants in New Zealand
have a combined output of $\sim$~1000MW.

To also investigate changes in periodicity we computed the continuous wavelet
transform of of $\dot{Q_i}$ using a complex Morlet wavelet (Figure
\ref{inference_result}b). We chose the wavelet transform over the more common
windowed Fourier transform for its ability to detect non-stationary signals
\cite{torrencePracticalGuideWavelet1998}. Figure \ref{inference_result}c shows
the global wavelet spectrum which is an average of the wavelet analysis in
Figure \ref{inference_result}b over the whole analysis period. The global
wavelet spectrum can be interpreted similarly to the Power Spectral Density.
Figure \ref{inference_result} shows that periodicity is dominated by periods
between 100 and 250 days. Which period is dominant at any one time changes
significantly and we will expand on this further in the following section.

\begin{figure}
 	\includegraphics[width=12 cm]{figures/time_frequency_analysis.png}  
    \caption{ a) Inference results for the heat input rate $\dot{Q_i}$ at RCL
             between 4 March 2016 and 1 February 2022. The solid line shows the
             marginal probability and the uncertainty is displayed as shaded
             region; b) Continuous wavelet transform of the time series shown in
             Panel a; c) Power spectral density of the heat input rate shown as a
             solid line with uncertainty displayed as shaded region.}
\label{inference_result}
\end{figure}

\section{Discussion and conclusion}
Inferring the heat input rate into a crater lake is an important part in
estimating the total energy budget of a volcanic system \cite{Brown1989a}.
Similar to a volcano's gas budget, changes in the heat input rate can either
indicate changes in the magmatic heat source or in the hydrothermal system
above.

We would expect a significant change in the heat source to cause a change in
slope of the cumulative heat input rate (Figure \ref{cumulative}). The fact that
we do not observe such changes may indicate a more or less constant supply of
heat during the studied period. Experiments on analog models of hydrothermal
systems suggest that changes in the heat source and the hydrothermal system
should also change the length of heating cycles
\cite{vandemeulebrouckAnalogueModelingInstabilities2005,fitzgerald1997TwentySecondWorkshop}.
Figure \ref{inference_result}b shows the heating cycles before mid-2017 and
after mid-2019 were dominated by longer periods. From mid-2017 to mid-2019
cycles were mostly of shorter periods which may indicate a change in the
volcanic system during this time.

\begin{figure}
 	\includegraphics[width=12 cm]{figures/cumulative_heat.png}  
    \caption{Cumulative heat input into RCL (solid line) and its uncertainty 
             (shaded region) calculated for the time-series in
             Figure \ref{inference_result} a.} 
\label{cumulative}
\end{figure}

To further explore hypotheses like this, the next logical step is to include
numerical models of magma degassing and hydrothermal heat and fluid flow. While
beyond the scope of this study, our inversion method is well suited for such
more complex models as its runtime scales linearly with the number of
parameters. Several studies have demonstrated the value of non-linear Kalman
Filters for inverse problems with non-linear, high-dimensional numerical models
\cite[e.g.][]{White2018a,huangIteratedKalmanMethodology2022}.

Previous studies of RCL's mass and energy balance found that they required very
high values of enthalpy for the incoming steam to avoid violating physical
boundary conditions such as negative values for the inflow rate of meteoric
water ($\dot{M_s}$) \cite{hurstCraterLakeEnergy2015, Stevenson1992}. Their
explanation was that as rising hot fluids enter the lake, relatively cool lake
water flows back into the vent. This would lead to an overall lower addition of
mass to the lake and resulting in an enthalpy value that is double than expected
for saturated steam at hydrostatic pressures likely to be found at the bottom of
the lake. By including uncertainties in our observations, this explanation is
still a possibility but no longer a requirement to satisfy the physical
constraints. Numerical modelling of hydrothermal processes may also shed further
light on the heating processes and whether or not the lake and the vent form a
single or two separate convective systems.

In conclusion, the combination of the mass and energy model, developed in
previous studies \cite{hurstCraterLakeEnergy2015, Stevenson1992}, with the
probabilistic inference method we developed here provides a powerful method to
continuously infer the heat input rate into RCL. It allows us to combine
disparate data streams, quantify our beliefs in the results and thereby augment
our monitoring capabilities by keeping a finger on Mt. Ruapehu's energy budget.

\appendix
\section{Surface losses}\label{A}
\subsection{long-wavelength radiation}
The net loss through long-wavelength radiation can be written as:
\begin{equation}
    Q_{rad} = A\sigma(\epsilon_w T_s^4 - \epsilon_a T_a^4)
\end{equation}

where
\begin{table}[h!]
\begin{tabular}{cp{8cm}}
    A & is the surface of the lake (in $m^2$) \\
    $\sigma$ & is the Stefan-Boltzmann coefficient ($5.67e^{-8} \frac{W}{m^2 K^4}$) \\
    $\epsilon_w$ & is the lake emissivity \\
    $T_s$ & is the water surface temperature (here assumed to be 1 $^{\circ}C$ less than the
    measured temperature) \\
    $\epsilon_a$ & is the effective atmospheric emissivity \\
    $T_a$ & is the effective air temperature (here assumed to be 0.9 $^{\circ}C$) \\ 
\end{tabular}
\end{table}

\subsection{Evaporation}
As proposed by \cite{hurstCraterLakeEnergy2015} we use the equation of
\cite{adamsEvaporationHeatedWater1990} but replace the term for forced
convection by that of \cite{sartoriCriticalReviewEquations2000}:

\begin{equation}
    Q_{evap} = [(2.2(T_{sv} - T_{av})^{1/3}(e_s - e_a))^2 + 
    (L(0.00407u^{0.8}X_0^{-0.2})(e_s - e_a)/P_a)^2]^{1/2}
\end{equation}

where
\begin{table}[h!]
\begin{tabular}{cp{8cm}}
    $T_{sv}$, $T_{av}$ & are the virtual surface and air temperatures respectively.
    This correction accounts for the extra buoyancy of water vapour compared to air
    (see Equation \ref{vtemp} for details) \\
    $e_s$, $e_a$ & are the saturation vapour pressure and the ambient air vapour
    pressure respectively \\
    $L$ & is the latent heat of evaporation \\
    $u$ & is the wind speed \\
    $X_0$ & is the fetch, or characteristic length, of the lake\\
    $P_a$ & is the atmospheric pressure \\
\end{tabular}
\end{table}

\begin{equation}\label{vtemp}
T_{xv} = \frac{T_x}{(1-0.378e_x/P_a)} \qquad \mathrm{for}\quad x = s\ \mathrm{or}\ a
\end{equation}

\subsection{Sensible heat}
This describes the heat loss due to conduction and convection above a hot lake
and according to \cite{Stevenson1992} the ratio between sensible heat,
$Q_{sh}$, and heat loss due to evaporation, $Q_{evap}$ can be written as:

\begin{equation}
    \frac{Q_{sh}}{Q_{evap}}=R_{sh}=\frac{\rho c_a}{L} \frac{(T_s - T_a)}{(e_s - e_a)}
    \underbrace{\left [ \frac{(T_s-T_a)}{(T_{sv} - T_{av})} \right ]^{1/3}}_A 
\end{equation}

where

\begin{table}[h!]
\begin{tabular}{cp{8cm}}
    $\rho$ & is the air density \\
    $c_a$ & is the specific heat of air\\
\end{tabular}
\end{table}

We decided to ignore term A as it tends to be very close to 1.

\subsection{Putting it all together}
The final equation to compute surface heat losses is then:
\begin{equation}
    Q_e = Q_{rad} + Q_{evap}(1+R_{sh})
\end{equation}

and the mass-loss due to evaporation is:
\begin{equation}
    M_e = Q_{evap}/L
\end{equation}

%\nocite{oreg,schn,pond,smith,marg,hunn,advi,koha,mouse}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
    Text for this section \ldots

\section*{Acknowledgements}
  Text for this section \ldots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{zotero}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

%% \section*{Figures}
%%   \begin{figure}[h!]
%%   \caption{\csentence{Sample figure title.}
%%       A short description of the figure content
%%       should go here.}
%%       \end{figure}
%% 
%% \begin{figure}[h!]
%%   \caption{\csentence{Sample figure title.}
%%       Figure legend text.}
%%       \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
%% \section*{Tables}
%% \begin{table}[h!]
%% \caption{Sample table title. This is where the description of the table should go.}
%%       \begin{tabular}{cccc}
%%         \hline
%%            & B1  &B2   & B3\\ \hline
%%         A1 & 0.1 & 0.2 & 0.3\\
%%         A2 & ... & ..  & .\\
%%         A3 & ..  & .   & .\\ \hline
%%       \end{tabular}
%% \end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% \section*{Additional Files}
%%   \subsection*{Additional file 1 --- Sample additional file title}
%%     Additional file descriptions text (including details of how to
%%     view the file, if it is in a non-standard format or the file extension).  This might
%%     refer to a multi-page table or a figure.
%% 
%%   \subsection*{Additional file 2 --- Sample additional file title}
%%     Additional file descriptions text.

\end{backmatter}
\end{document}